The Proofs

### 2023-11-13 proofs/main.tex
Fix margins

\begin{algorithm}
  \hfill\\
  \ottdefnATNInfLabeled{}
  \ottdefnATPInfLabeled{}
  \ottdefnATSpinInfLabeled{}
\end{algorithm}

@@

First, we define the notion of singularity on constraint entries. 
$[[scE singular with iP]]$ and $[[scE singular with iN]]$
are considered partial functions taking a constraint entry $[[scE]]$
and returning the type satisfying $[[scE]]$ if such a type is unique.

>>>>

\begin{algorithm}
  \hfill \\
  \ottdefnATPInfLabeled{}
  \hfill \\
  \ottdefnATNInfLabeled{}
  \hfill \\
  \ottdefnATSpinInfLabeled{}
\end{algorithm}

@@

First, we define the notion of singularity on constraint entries. 
$[[scE singular with iP]]$ and its negative counterpart 
are considered partial functions taking a constraint entry $[[scE]]$
and returning the type satisfying $[[scE]]$ if such a type is unique.

### 2023-11-13 proofs/tex/alg-pos-subtyping-lemmas.tex
Fix margins

                Notice that $[[β̂±]] \in [[dom(Θ)]]$, which is disjoint from $[[{α̂⁻*}]]$, 
                that is $[[β̂±]] \in [[dom(SC') \ {α̂⁻*}]]$. This way,
                $[[ [uσ'][â⁻*/nas]β̂± ]] = [[  [uσ']β̂± ]] = [[ [uσ]β̂± ]]$ by the definition 
                of $[[uσ']]$,
            \item If this variable is a regular variable $[[β±]] \notin [[nas]]$, then 
                $[[ [iNs/nas][uσ] β± ]] = [[ β± ]] $ and $[[ [uσ'][â⁻*/nas]β± ]] = [[ β± ]]$. 
            \item If this variable is a regular variable $[[αi⁻]] \in [[nas]]$, then 
                $[[ [iNs/nas][uσ] αi⁻ ]] = [[ iNi ]] = [[ nf(iNi) ]]$
                (the latter equality holds since $[[iNi]]$ is normalized)
                and $[[ [uσ'][â⁻*/nas]αi⁻ ]] = [[  [uσ']αî⁻ ]] = [[ nf(iNi) ]]$.
        \end{enumerate}
    \end{caseof}
\end{proof}

@@

            Now, let us show that $[[Γ, nbs ⊢ [us0]uP0 ≥ iQ]]$.
            To do that, we notice that $[[ [uσ0]uP0 ]] = [[ [uσ][σ][nas/â⁻*]uP0 ]]$:
            let us consider an arbitrary variable appearing freely in $[[uP0]]$:
            \begin{enumerate}
                \item if this variable is a algorithmic variable $[[αî⁻]] \in [[â⁻*]]$, then
                $[[ [uσ0]αî⁻ ]] = [[ [σ]αi⁻ ]]$ and 
                $[[ [uσ][σ][nas/â⁻*]αî⁻ ]] = [[ [uσ][σ]αi⁻ ]] = [[ [σ]αi⁻ ]]$,
                \item if this variable is a algorithmic variable $[[β̂±]] \in [[ uv(uP0) \ {â⁻*} ]] = [[uv(uP')]]$, then
                $[[ [uσ0]β̂± ]] = [[ [uσ]β̂± ]]$ and $[[ [uσ][σ][nas/â⁻*]β̂± ]] = [[ [uσ][σ]β̂± ]] = [[ [uσ]β̂± ]]$,
                \item if this variable is a regular variable from $[[fv(uP0)]]$, both substitutions do not change it:
                $[[ uσ0 ]]$, $[[ uσ ]]$ and $[[ nas / â⁻* ]]$ act on algorithmic variables, 
                and $[[σ]]$ is defined on $[[nas]]$, however, $[[{nas} ∩ fv(uP0) = ∅]]$.
            \end{enumerate}
            This way, $[[ [uσ0]uP0 ]] = [[ [uσ][σ][nas/â⁻*]uP0 ]] = [[ [uσ][σ]uP' ]]$,
            and thus, $[[ Γ, nbs ⊢ [uσ0]uP0 ≥ iQ' ]]$.
    \end{caseof}
\end{proof}

>>>>

                Notice that $[[β̂±]] \in [[dom(Θ)]]$, which is disjoint from $[[{α̂⁻*}]]$, 
                that is $[[β̂±]] \in [[dom(SC') \ {α̂⁻*}]]$. This way,
                $[[ [uσ'][â⁻*/nas]β̂± ]] = [[  [uσ']β̂± ]] = [[ [uσ]β̂± ]]$ by the definition 
                of $[[uσ']]$,
            \item If this variable is a regular variable $[[β±]] \notin [[nas]]$, then 
                $[[ [iNs/nas][uσ] β± ]] = [[ β± ]] $ and\\ $[[ [uσ'][â⁻*/nas]β± ]] = [[ β± ]]$. 
            \item If this variable is a regular variable $[[αi⁻]] \in [[nas]]$, then 
                $[[ [iNs/nas][uσ] αi⁻ ]] = [[ iNi ]] = [[ nf(iNi) ]]$
                (the latter equality holds since $[[iNi]]$ is normalized)
                and $[[ [uσ'][â⁻*/nas]αi⁻ ]] = [[  [uσ']αî⁻ ]] = [[ nf(iNi) ]]$.
        \end{enumerate}
    \end{caseof}
\end{proof}

@@

            Now, let us show that $[[Γ, nbs ⊢ [us0]uP0 ≥ iQ]]$.
            To do that, we notice that\\ $[[ [uσ0]uP0 ]] = [[ [uσ][σ][nas/â⁻*]uP0 ]]$:
            let us consider an arbitrary variable appearing freely in $[[uP0]]$:
            \begin{enumerate}
                \item if this variable is a algorithmic variable $[[αî⁻]] \in [[â⁻*]]$, then
                $[[ [uσ0]αî⁻ ]] = [[ [σ]αi⁻ ]]$ and\\
                $[[ [uσ][σ][nas/â⁻*]αî⁻ ]] = [[ [uσ][σ]αi⁻ ]] = [[ [σ]αi⁻ ]]$,
                \item if this variable is a algorithmic variable $[[β̂±]] \in [[ uv(uP0) \ {â⁻*} ]] = [[uv(uP')]]$, then
                $[[ [uσ0]β̂± ]] = [[ [uσ]β̂± ]]$ and $[[ [uσ][σ][nas/â⁻*]β̂± ]] = [[ [uσ][σ]β̂± ]] = [[ [uσ]β̂± ]]$,
                \item if this variable is a regular variable from $[[fv(uP0)]]$, both substitutions do not change it:
                $[[ uσ0 ]]$, $[[ uσ ]]$ and $[[ nas / â⁻* ]]$ act on algorithmic variables, 
                and $[[σ]]$ is defined on $[[nas]]$, however, $[[{nas} ∩ fv(uP0) = ∅]]$.
            \end{enumerate}
            This way, $[[ [uσ0]uP0 ]] = [[ [uσ][σ][nas/â⁻*]uP0 ]] = [[ [uσ][σ]uP' ]]$,
            and thus, $[[ Γ, nbs ⊢ [uσ0]uP0 ≥ iQ' ]]$.
    \end{caseof}
\end{proof}

### 2023-11-13 proofs/tex/norm-lemmas.tex
Fix margins

      $[[ [nf(σ)] nf(iN) ]] = [[ [nf(σ)] nf(na) ]] = [[ [nf(σ)] na ]] $.
    \item $[[iP]]$ = $[[pa]]$ \\
      Similar to \cref{case:norm-subst-distr-var}.
   \item If the type is formed by $[[→]]$, $[[↑]]$, or $[[↓]]$, 
     the required equality follows from the congruence of the normalization and
     substitution, and the induction hypothesis.
     For example, if $[[iN]] = [[iP → iM]]$ then \\
     $\begin{aligned}[t]
        [[nf([σ] iN)]] &= [[ nf([σ] (iP → iM)) ]] \\
                        &= [[ nf([σ]iP → [σ]iM) ]]
                        && \text{By the congruence of substitution} \\
                        &= [[ nf([σ]iP) → nf([σ]iM) ]]
                        && \text{By the congruence of normalization, i.e. \ruleref{\ottdruleNrmArrowLabel}} \\
                        &= [[ [nf(σ)]nf(iP) → [nf(σ)]nf(iM) ]]
                        && \text{By the induction hypothesis} \\
                        &= [[ [nf(σ)](nf(iP) → nf(iM)) ]]
                        && \text{By the congruence of substitution} \\
                        &= [[ [nf(σ)]nf(iP → iM) ]]
                        && \text{By the congruence of normalization} \\
                        &= [[ [nf(σ)]nf(iN) ]]
      \end{aligned}$ \\
    \item $[[iN]] = [[∀ pas.iM]]$ \label{case:norm-subst-commute} \\
      $\begin{aligned}[t]
          [[ [nf(σ)] nf(iN) ]] &= [[ [nf(σ)] nf(∀pas.iM)]] \\
                            &= [[ [nf(σ)] ∀pas'.nf(iM) ]]
                            && \text {Where $[[pas']] = [[ ord {pas} in nf(iM)]]
                               = [[ord {pas} in iM]]$
                               (the latter is by
                               \cref{corollary:normalization-ord})}\\
        \end{aligned}$ \\

      $\begin{aligned}[t]
         [[ nf([σ]iN) ]] &= [[ nf([σ] ∀pas.iM)]] \\
                          &= [[ nf(∀pas.[σ]iM) ]]
                          && \text{Assuming $[[{pas} ∩ Γ1]] = \emptyset$
                             and $[[{pas} ∩ Γ2]] = \emptyset$}\\
                          &= [[ ∀pbs.nf([σ]iM) ]]
                          && \text {Where $[[pbs]] = [[ord {pas} in nf([σ]iM)]]
                             = [[ord {pas} in [σ]iM]]$ (the latter is by
                             \cref{corollary:normalization-ord})}\\
                          &= [[ ∀pas'.nf([σ]iM) ]]
                          && \text{By \cref{lemma:ord-sigma}, $[[pbs]] = [[pas']]$
                             since $[[{pas}]]$ is disjoint with $[[Γ1]]$ and
                             $[[Γ2]]$}\\
                          &= [[ ∀pas'.[nf(σ)]nf(iM) ]]
                          && \text {By the induction hypothesis}\\
         \end{aligned}$ \\

@@

    \item[(ii)] $\begin{aligned}[t] [[ [mu]pbs' ]] &= [[ [mu] ord {pbs} in nf(iM) ]]
                                                  && \text{by the definition of $[[pbs']]$ } \\
                                                  &= [[ [mu] ord ({pbs} ∩ fv iM) in nf(iM) ]]
                                                  && \text{from \cref{lemma:fv-nf,corollary:ord-weakening} } \\
                                                  &= [[ ord [mu] ({pbs} ∩ fv iM) in [mu] nf(iM) ]]
                                                  && \text{by \cref{lemma:distr-mu-ord}, because
                                                     $[[{pas} ∩ fv iN]] \cap [[fv nf(iM)]] \subseteq [[{pas} ∩ fv iM ]]
                                                     = \emptyset$}\\
                                                  &
                                                  && \text{and $[[{pas} ∩ fv iN]] \cap [[({pbs} ∩ fv iM)]] \subseteq
                                                     [[{pas} ∩ fv iM]] = \emptyset$} \\
                                                  &= [[ ord [mu] ({pbs} ∩ fv iM) in nf(iN) ]]
                                                  && \text{since $[[ [mu] nf(iM) ]] = [[nf(iN)]]$ is proved } \\
                                                  &= [[ ord ({pas} ∩ fv iN) in nf(iN) ]]
                                                  && \text{because $\mu$ is a bijection between
                                                     $[[{pas} ∩ fv iN]]$ and $[[{pbs} ∩ fv iM]]$} \\
                                                  &= [[ ord {pas} in nf(iN) ]]
                                                  && \text{from \cref{lemma:fv-nf,corollary:ord-weakening} } \\
                                                  &= [[ pas' ]]
                                                  && \text{by the definition of $[[pas']]$} \\
      \end{aligned}$
    \end{enumerate}
  \item {\nameref{\ottdruleEOneExistsLabel}} Same as for \cref{case:ord-completeness:forall}.
  \item Other rules are congruent, and thus, proved by the corresponding congruent alpha-equivalence rule,
    which is applicable by the induction hypothesis. 
  \end{caseof}
\end{proof}

@@

                                           && \text{by the substitution

@@

      [[ [σ](∀pas.iM) ]] \text{ is normal} &\iff [[ (∀pas.[σ]iM) ]] \text{ is normal}

>>>>

      $[[ [nf(σ)] nf(iN) ]] = [[ [nf(σ)] nf(na) ]] = [[ [nf(σ)] na ]] $.
    \item $[[iP]]$ = $[[pa]]$ \\
      Similar to \cref{case:norm-subst-distr-var}.
   \item If the type is formed by $[[→]]$, $[[↑]]$, or $[[↓]]$, 
     the required equality follows from the congruence of the normalization and
     substitution and the induction hypothesis.
     For example, if $[[iN]] = [[iP → iM]]$ then \\
     $\begin{aligned}[t]
        [[nf([σ] iN)]] &= [[ nf([σ] (iP → iM)) ]] \\
                        &= [[ nf([σ]iP → [σ]iM) ]]
                        && \text{By congruence of substitution} \\
                        &= [[ nf([σ]iP) → nf([σ]iM) ]]
                        && \text{By congruence of normalization} \\
                        &= [[ [nf(σ)]nf(iP) → [nf(σ)]nf(iM) ]]
                        && \text{By induction hypothesis} \\
                        &= [[ [nf(σ)](nf(iP) → nf(iM)) ]]
                        && \text{By congruence of substitution} \\
                        &= [[ [nf(σ)]nf(iP → iM) ]]
                        && \text{By congruence of normalization} \\
                        &= [[ [nf(σ)]nf(iN) ]]
      \end{aligned}$ \\
    \item $[[iN]] = [[∀ pas.iM]]$ \label{case:norm-subst-commute} \\
      $\begin{aligned}[t]
          [[ [nf(σ)] nf(iN) ]] &= [[ [nf(σ)] nf(∀pas.iM)]] \\
                            &= [[ [nf(σ)] ∀pas'.nf(iM) ]]
                            && \text {Where $[[pas']] = [[ ord {pas} in nf(iM)]]
                               = [[ord {pas} in iM]]$}\\
                           & && 
                               \text{(the latter is by
                               \cref{corollary:normalization-ord})}\\
        \end{aligned}$ \\

      $\begin{aligned}[t]
         [[ nf([σ]iN) ]] &= [[ nf([σ] ∀pas.iM)]] \\
                          &= [[ nf(∀pas.[σ]iM) ]]
                          && \text{Assuming $[[{pas} ∩ Γ1]] = \emptyset$
                             and $[[{pas} ∩ Γ2]] = \emptyset$}\\
                          &= [[ ∀pbs.nf([σ]iM) ]]
                          && \text {Where $[[pbs]] = [[ord {pas} in nf([σ]iM)]]
                             = [[ord {pas} in [σ]iM]]$}\\
                          & && 
                             \text{(the latter is by \cref{corollary:normalization-ord})}\\
                          &= [[ ∀pas'.nf([σ]iM) ]]
                          && \text{By \cref{lemma:ord-sigma},}\\
                          & && \text{$[[pbs]] = [[pas']]$
                             since $[[{pas}]]$ is disjoint with $[[Γ1]]$ and
                             $[[Γ2]]$}\\
                          &= [[ ∀pas'.[nf(σ)]nf(iM) ]]
                          && \text {By the induction hypothesis}\\
         \end{aligned}$ \\

@@

    \item[(ii)] $\begin{aligned}[t] [[ [mu]pbs' ]] &= [[ [mu] ord {pbs} in nf(iM) ]]
                                                  && \text{by the definition of $[[pbs']]$ } \\
                                                  &= [[ [mu] ord ({pbs} ∩ fv iM) in nf(iM) ]]
                                                  && \text{from \cref{lemma:fv-nf,corollary:ord-weakening} } \\
                                                  &= [[ ord [mu] ({pbs} ∩ fv iM) in [mu] nf(iM) ]]
                                                  && \text{by \cref{lemma:distr-mu-ord}, because}\\
                                                  & && \text{$[[{pas} ∩ fv iN]] \cap [[fv nf(iM)]] \subseteq [[{pas} ∩ fv iM ]]
                                                     = \emptyset$}\\
                                                  &
                                                  && \text{$[[{pas} ∩ fv iN]] \cap [[({pbs} ∩ fv iM)]] \subseteq
                                                     [[{pas} ∩ fv iM]] = \emptyset$} \\
                                                  &= [[ ord [mu] ({pbs} ∩ fv iM) in nf(iN) ]]
                                                  && \text{since $[[ [mu] nf(iM) ]] = [[nf(iN)]]$ is proved } \\
                                                  &= [[ ord ({pas} ∩ fv iN) in nf(iN) ]]
                                                  && \text{because $\mu$ is a bijection between}\\
                                                  & && \text{$[[{pas} ∩ fv iN]]$ and $[[{pbs} ∩ fv iM]]$} \\
                                                  &= [[ ord {pas} in nf(iN) ]]
                                                  && \text{from \cref{lemma:fv-nf,corollary:ord-weakening} } \\
                                                  &= [[ pas' ]]
                                                  && \text{by the definition of $[[pas']]$} \\
      \end{aligned}$
    \end{enumerate}
  \item {\nameref{\ottdruleEOneExistsLabel}} Same as for \cref{case:ord-completeness:forall}.
  \item Other rules are congruent, and thus, proved by the corresponding congruent alpha-equivalence rule,
    which is applicable by the induction hypothesis. 
  \end{caseof}
\end{proof}

@@

                                           && \text{by substitution

@@

      &[[ [σ](∀pas.iM) ]] \text{ is normal} \\
       &\iff [[ (∀pas.[σ]iM) ]] \text{ is normal}

### 2023-11-13 proofs/tex/ord-lemmas.tex
Fix margins

      Let us show that that $A = B$ and $X = Y$, so the results are equal. 
      By the induction hypothesis and set properties,
      $[[ ord (varset ∩ (fv iP ∪ fv iN)) in iP ]] = 
       [[ ord (varset ∩ (fv iP ∪ fv iN)) ∩ fv(iP) in iP ]] = 
       [[ ord varset ∩ fv(iP) in iP ]] = 
       [[ ord varset in iP ]]$.
      Analogously, 
      $[[ ord (varset ∩ (fv iP ∪ fv iN)) in iN ]] = 
       [[ ord varset in iN ]]$.
    \item If the term has shape $[[∀pas.iN]]$,
      we can assume that $[[pas]]$ is disjoint
      from $[[varset]]$,
      since we operate on alpha-equivalence classes.
      Then using the induction hypothesis,
      set properties and \ruleref{\ottdruleOForallLabel}: 
      $[[ord varset ∩ (fv(∀pas.iN)) in ∀pas.iN]] =
       [[ord varset ∩ (fv(iN) \ {pas}) in iN]] =
       [[ord varset ∩ (fv(iN) \ {pas}) ∩ fv(iN) in iN]] =
       [[ord varset ∩ fv(iN) in iN]] =
       [[ord varset in iN]]$.
  \end{caseof}
\end{proof}

@@

\begin{definition} [Collision-free bijection]
  We say that a bijection $\mu : A \leftrightarrow B$ between sets of
  variables is \textbf{collision free on sets} $P$ and $Q$ if and only if
  \begin{enumerate}
    \item $\mu(P \cap A) \cap Q = \emptyset$
    \item $\mu(Q \cap A) \cap P = \emptyset$
  \end{enumerate}
\end{definition}

\begin{observation}
  Suppose that $\mu : A \leftrightarrow B$ is a bijection between two sets of variables,
  and $\mu$ is collision free on $P$ and $Q$.
  Then $\mu(P \cap Q) = \mu(P) \cap \mu(Q)$.
\end{observation}

\lemDistrMuOrd*
\begin{proof}
  Mutual induction on $[[iN]]$ and $[[iP]]$.
  \begin{caseof}
  \item $[[iN]]$ = $[[na]]$ \label{case:distr-mu-ord:var} \\
    let us consider four cases:
    \begin{caseof}
    \item $[[na]] \in A$ and $[[na]] \in [[varset]]$\\ Then
      $
      \begin{aligned}[t] [[ [mu] (ord varset in iN) ]] &= [[ [mu] (ord varset in na)]] \\
                                                             &= [[ [mu] na ]]
                                                             && \text{by \ruleref{\ottdruleOPVarInLabel}}\\
                                                             &= [[nb]]
                                                             && \text{for some $[[nb]] \in B$ (notice that $[[nb]] \in [[ [mu]varset ]]$)} \\
                                                             &= [[ ord [mu]varset in nb ]]
                                                             && \text{by \ruleref{\ottdruleOPVarInLabel},
                                                                because $[[nb]] \in [[ [mu]varset ]]$} \\
                                                             &= [[ord [mu] varset in [mu] na ]]
       \end{aligned}
       $
     \item $[[na]] \notin A$ and $[[na]] \notin [[varset]]$\\
       Notice that
       $[[ [mu] (ord varset in iN) ]] = [[ [mu] (ord varset in na)]] = [[·]]$ by
       \ruleref{\ottdruleOPVarNInLabel}.
       On the other hand, $[[ ord [mu] varset in [mu] na = ord [mu] varset
       in na ]] = [[·]]$ The latter equality is from
       \ruleref{\ottdruleOPVarNInLabel}, because
       $[[mu]]$ is collision free on $[[varset]]$ and $[[fv iN]]$, so
       $[[fv iN]] \ni [[na]] \notin [[mu]](A \cap [[varset]]) \cup
       [[varset]] \supseteq [[ [mu] varset ]]$.
     \item $[[na]] \in A$ but $[[na]] \notin [[varset]]$\\ Then
       $[[ [mu] (ord varset in iN) ]] = [[ [mu] (ord varset in na)]] = [[·]]$
       by \ruleref{\ottdruleOPVarNInLabel}.
       To prove that $[[ ord [mu] varset in [mu] na ]] = [[·]]$, we apply
       \ruleref{\ottdruleOPVarNInLabel}. Let us show that
       $[[ [mu] na ]] \notin [[ [mu] varset ]]$.
       Since $[[ [mu] na ]] = [[mu]]([[na]])$ and
       $[[ [mu] varset ]] \subseteq [[mu]](A \cap [[varset]]) \cup [[varset]]$,
       it suffices to prove 
       $[[mu]]([[na]]) \notin [[mu]](A \cap [[varset]]) \cup [[varset]]$.

       \begin{enumerate}
       \item[(i)] If there is an element $x \in A \cap [[varset]]$ such that
         $[[mu]] x = [[mu]] [[na]]$, then $x = [[na]]$ by bijectivity of
         $[[mu]]$, which contradicts with $[[na]] \notin [[varset]]$. This way, 
         $[[mu]]([[na]]) \notin [[mu]](A \cap [[varset]])$.
       \item[(ii)]
         Since $[[mu]]$ is collision free on $[[varset]]$ and $[[fv iN]]$,
         $[[mu]] (A \cap [[fv iN]]) \ni [[mu]]([[na]]) \notin [[varset]]$.
       \end{enumerate}

@@

   \item $[[iN]] = [[iP → iM]]$  \\
     $\begin{aligned}[t]
        [[ [mu] (ord varset in iN) ]] &= [[ [mu] (ord varset in iP → iM) ]] \\
                                      &= [[ [mu] (ordVars1, (ordVars2 \ {ordVars1})) ]]
                                      && \text{where } [[ord varset in iP = ordVars1]] \text{ and } [[ord varset in iM = ordVars2]] \\
                                      &= [[ [mu] ordVars1, [mu](ordVars2 \ {ordVars1}) ]] \\
                                      &= [[ [mu] ordVars1, ([mu]ordVars2 \ [mu]{ordVars1}) ]]
                                      && \text{by induction on $[[ordVars2]]$;
                                         the inductive step is similar to \cref{case:distr-mu-ord:var}.
                                         Notice that $[[mu]]$ is} \\
                                      & && \text{collision free on $[[{ordVars1}]]$ and $[[{ordVars2}]]$
                                           since
                                           $[[{ordVars1}]] \subseteq [[varset]]$ and
                                           $[[{ordVars2}]] \subseteq [[fv iN]]$ }\\
                                      &= [[ [mu] ordVars1, ([mu]ordVars2 \ {[mu]ordVars1}) ]]
      \end{aligned}$ \\
    $\begin{aligned}[t]
       [[  (ord [mu] varset in [mu]iN) ]] &= [[ (ord [mu] varset in [mu]iP → [mu]iM) ]] \\
                                     &= [[ (ordVarsb1, (ordVarsb2 \ {ordVarsb1})) ]]
                                     && \text{where } [[ord [mu] varset in [mu] iP = ordVarsb1]] \text{ and } [[ord [mu] varset in [mu] iM = ordVarsb2]] \\
                                          & && \text{then by the induction
                                               hypothesis,
                                               $[[ordVarsb1]] = [[ [mu] ordVars1 ]]$,
                                               $[[ordVarsb2]] = [[ [mu] ordVars2 ]]$,
                                               }\\
                                     &= [[ [mu] ordVars1, ([mu]ordVars2 \ {[mu]ordVars1}) ]]
     \end{aligned}$

>>>>

      Let us show that $A = B$ and $X = Y$, so the results are equal. 
      By the induction hypothesis and set properties,
      $[[ ord (varset ∩ (fv iP ∪ fv iN)) in iP ]] = 
       [[ ord (varset ∩ (fv iP ∪ fv iN)) ∩ fv(iP) in iP ]] = 
       [[ ord varset ∩ fv(iP) in iP ]] = 
       [[ ord varset in iP ]]$.
      Analogously, 
      $[[ ord (varset ∩ (fv iP ∪ fv iN)) in iN ]]$ $=$
      \\ $[[ ord varset in iN ]]$.
    \item If the term has shape $[[∀pas.iN]]$,
      we can assume that $[[pas]]$ is disjoint
      from $[[varset]]$,
      since we operate on alpha-equivalence classes.
      Then using the induction hypothesis,
      set properties and \ruleref{\ottdruleOForallLabel}: 
      $[[ord varset ∩ (fv(∀pas.iN)) in ∀pas.iN]] =
       [[ord varset ∩ (fv(iN) \ {pas}) in iN]] =
       [[ord varset ∩ (fv(iN) \ {pas}) ∩ fv(iN) in iN]] =
       [[ord varset ∩ fv(iN) in iN]] =
       [[ord varset in iN]]$.
  \end{caseof}
\end{proof}

@@

\begin{definition} [Collision-free bijection]
  We say that a bijection $\mu : A \leftrightarrow B$ between sets of
  variables is \textbf{collision-free on sets} $P$ and $Q$ if and only if
  \begin{enumerate}
    \item $\mu(P \cap A) \cap Q = \emptyset$
    \item $\mu(Q \cap A) \cap P = \emptyset$
  \end{enumerate}
\end{definition}

\begin{observation}
  Suppose that $\mu : A \leftrightarrow B$ is a bijection between two sets of variables,
  and $\mu$ is collision-free on $P$ and $Q$.
  Then $\mu(P \cap Q) = \mu(P) \cap \mu(Q)$.
\end{observation}

\lemDistrMuOrd*
\begin{proof}
  Mutual induction on $[[iN]]$ and $[[iP]]$.
  \begin{caseof}
  \item $[[iN]]$ = $[[na]]$ \label{case:distr-mu-ord:var} \\
    let us consider four cases:
    \begin{caseof}
    \item $[[na]] \in A$ and $[[na]] \in [[varset]]$\\ Then
    \\
      $
      \begin{aligned}[t] [[ [mu] (ord varset in iN) ]] &= [[ [mu] (ord varset in na)]] \\
                                                             &= [[ [mu] na ]]
                                                             && \text{by \ruleref{\ottdruleOPVarInLabel}}\\
                                                             &= [[nb]]
                                                             && \text{for some $[[nb]] \in B$ (notice $[[nb]] \in [[ [mu]varset ]]$)} \\
                                                             &= [[ ord [mu]varset in nb ]]
                                                             && \text{by \ruleref{\ottdruleOPVarInLabel},
                                                                because $[[nb]] \in [[ [mu]varset ]]$} \\
                                                             &= [[ord [mu] varset in [mu] na ]]
       \end{aligned}
       $
     \item $[[na]] \notin A$ and $[[na]] \notin [[varset]]$\\
       Notice that
       $[[ [mu] (ord varset in iN) ]] = [[ [mu] (ord varset in na)]] = [[·]]$ by
       \ruleref{\ottdruleOPVarNInLabel}.
       On the other hand, $[[ ord [mu] varset in [mu] na = ord [mu] varset
       in na ]] = [[·]]$ The latter equality is from
       \ruleref{\ottdruleOPVarNInLabel}, because
       $[[mu]]$ is collision-free on $[[varset]]$ and $[[fv iN]]$, so
       $[[fv iN]] \ni [[na]] \notin [[mu]](A \cap [[varset]]) \cup
       [[varset]] \supseteq [[ [mu] varset ]]$.
     \item $[[na]] \in A$ but $[[na]] \notin [[varset]]$\\ Then
       $[[ [mu] (ord varset in iN) ]] = [[ [mu] (ord varset in na)]] = [[·]]$
       by \ruleref{\ottdruleOPVarNInLabel}.
       To prove that\\ $[[ ord [mu] varset in [mu] na ]] = [[·]]$, we apply
       \ruleref{\ottdruleOPVarNInLabel}. Let us show that
       $[[ [mu] na ]] \notin [[ [mu] varset ]]$.
       Since $[[ [mu] na ]] = [[mu]]([[na]])$ and
       $[[ [mu] varset ]] \subseteq [[mu]](A \cap [[varset]]) \cup [[varset]]$,
       it suffices to prove 
       $[[mu]]([[na]]) \notin [[mu]](A \cap [[varset]]) \cup [[varset]]$.

       \begin{enumerate}
       \item[(i)] If there is an element $x \in A \cap [[varset]]$ such that
         $[[mu]] x = [[mu]] [[na]]$, then $x = [[na]]$ by bijectivity of
         $[[mu]]$, which contradicts with $[[na]] \notin [[varset]]$. This way, 
         $[[mu]]([[na]]) \notin [[mu]](A \cap [[varset]])$.
       \item[(ii)]
         Since $[[mu]]$ is collision-free on $[[varset]]$ and $[[fv iN]]$,
         $[[mu]] (A \cap [[fv iN]]) \ni [[mu]]([[na]]) \notin [[varset]]$.
       \end{enumerate}

@@

   \item $[[iN]] = [[iP → iM]]$  \\
     $\begin{aligned}[t]
        & [[ [mu] (ord varset in iN) ]] \\ 
        &= [[ [mu] (ord varset in iP → iM) ]] \\
        &= [[ [mu] (ordVars1, (ordVars2 \ {ordVars1})) ]]
          && \text{where } [[ord varset in iP = ordVars1]] \text{ and } [[ord varset in iM = ordVars2]] \\
        &= [[ [mu] ordVars1, [mu](ordVars2 \ {ordVars1}) ]] \\
        &= [[ [mu] ordVars1, ([mu]ordVars2 \ [mu]{ordVars1}) ]]
          && \text{by induction on $[[ordVars2]]$;
                  the inductive step is similar to 
                  \cref{case:distr-mu-ord:var}.}\\
        & && \text{Notice that $[[mu]]$ is collision free on $[[{ordVars1}]]$ and $[[{ordVars2}]]$} \\
        & && \text{since
          $[[{ordVars1}]] \subseteq [[varset]]$ and
          $[[{ordVars2}]] \subseteq [[fv iN]]$ }\\
          &= [[ [mu] ordVars1, ([mu]ordVars2 \ {[mu]ordVars1}) ]]
      \end{aligned}$ \\
      \hfill\\
      On the other hand, \\
    $\begin{aligned}[t]
       & [[  ord [mu] varset in [mu]iN ]] \\ 
       &= [[ ord [mu] varset in [mu]iP → [mu]iM ]] \\
       &= [[ ordVarsb1, (ordVarsb2 \ {ordVarsb1}) ]]
       && \text{where } [[ord [mu] varset in [mu] iP = ordVarsb1]] \text{ and }  \\
       & && [[ord [mu] varset in [mu] iM = ordVarsb2]]; 
            \text{ then by the induction}\\
       & && \text{hypothesis, $[[ordVarsb1]] = [[ [mu] ordVars1 ]]$,
                   $[[ordVarsb2]] = [[ [mu] ordVars2 ]]$}\\
       &= [[ [mu] ordVars1, ([mu]ordVars2 \ {[mu]ordVars1}) ]]
     \end{aligned}$

### 2023-11-13 proofs/tex/singularity-lemmas.tex
Fix margins

            $
            \begin{aligned}[h]
            [[Γ ⊢ iP' ≥ ∃nbs.↓iN1]] &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbs'.↓nf(iN1)]] \text{ where } [[ord {nbs} in iN' = nbs']] 
                                   && \text{by \cref{corollary:nf-pres-subt}} \\
                                   &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbs'.↓nf(nbj)]]  
                                   && \text{by \cref{lemma:normalization-completeness}}\\
                                   &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbs'.↓nbn]]  
                                   && \text{by definition of normalization}\\
                                   &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbj.↓nbj]]  
                                   && \text{since $[[ord {nbs} in nf(iN1)]] = [[nbj]]$}\\
                                   &\Rightarrow [[Γ, nbj ⊢ nf(iP') ≥ ↓nbj]] \text { and } [[nbj ∉ fv(nf(iP'))]]
                                   && \text{by \cref{lemma:shape-supertypes-norm}}\\
            \end{aligned}
            $

@@

                            Then $[[scE singular with β⁺]]$, which holds by
                            \ruleref{\ottdruleSINGSupVarLabel},
                            can be rewritten as $[[scE singular with nf(iP)]]$.

>>>>

            $
            \begin{aligned}
            [[Γ ⊢ iP' ≥ ∃nbs.↓iN1]] &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbs'.↓nf(iN1)]]\\
                                    &\phantom{\Rightarrow} \text{(where $[[ord {nbs} in iN' = nbs']]$)}
                                    && \text{by \cref{corollary:nf-pres-subt}} \\
                                    &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbs'.↓nf(nbj)]]  
                                    && \text{by \cref{lemma:normalization-completeness}}\\
                                    &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbs'.↓nbn]]  
                                    && \text{by definition of normalization}\\
                                    &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbj.↓nbj]]  
                                    && \text{since $[[ord {nbs} in nf(iN1)]] = [[nbj]]$}\\
                                    &\Rightarrow [[Γ, nbj ⊢ nf(iP') ≥ ↓nbj]] \\ 
                                    &\phantom{\Rightarrow} \text { and } [[nbj ∉ fv(nf(iP'))]]
                                    && \text{by \cref{lemma:shape-supertypes-norm}}\\
            \end{aligned}
            $

@@

                            Then $[[scE singular with β⁺]]$, which holds by
                            \ruleref{\ottdruleSINGSupVarLabel},
                            can be rewritten as\\ $[[scE singular with nf(iP)]]$.
